Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-04 PandaOverview.tex
Space-saving section 2

PanDA is a Workload Management System (WMS) designed to support the execution of
distributed workloads and workflows via pilots. WMS is middleware for
discovering and selecting resources, submitting tasks of workloads and
workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
abstraction that enables multi-level scheduling by decoupling resource
acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
implemented, a pilot is scheduled on a site and, once active, tasks are
scheduled to the pilot, not to the site's scheduler.

Pilot-capable WMS enable high throughput of tasks execution while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data. The design of PanDA WMS
started in 2005 to support ATLAS. PanDA went into
production for the LHC Run 1 on 2009, and was then extended with new subsystems
for Run 2 in 2015.

@@

PanDA's application model assumes tasks, workloads and workflows. Tasks
represent a set of operations performed on data sets stored in one or more input
files. Tasks are decomposed into jobs, where each job represents the task's set
of operations and a partition of the task's data set. Since 2005, a certain
amount of parallelism has been progressively introduced for job
execution~\cite{crooks2012multi} but, so far, HEP applications have not required
MPI. Jobs are supposed to be relatively self-contained, capable of setting up
their own execution environment or having a minimal set of common dependences.

PanDA's usage model is based on multitenancy of resources and the support of at
least two types of HEP users: individual researchers and groups executing so
called `production' workflows. Users are free to submit tasks and workflows to
the PanDA WMS at any point in time, directly or via dedicated application
frameworks. PanDA's security model is based on separation between
authentication, authorization and accounting for both single users and group of
users. Both authentication and authorization are based on digital certificates
and the X.509 standard and on the virtual organization (VO) abstraction.

@@

PanDA's design supports provenance and traceability for both jobs and data.
Attributes enable provenance by linking jobs and data items, providing
information like ownership or project affiliation. States enable traceability by
providing information about the stage of the execution in which each job or data
item is or has been. Some attributes are assumed to be immutable across
execution and jobs and data items are assumed to be always in one and only one
state.

@@

The implementation of PanDA consists of several interconnected subsystems, most
of them built from off-the-shelf and Open Source components. Subsystems
communicate via dedicated API or HTTP messaging, and each subsystem is
implemented by one or more modules. Databases are used to store eventful
entities and entities' description like tasks, jobs and % events
input/output data, and to store information about sites,
resources, logs, and accounting.

@@

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During LHC Run 1, PanDA
required users or applications to perform a static conversion between tasks and
jobs: tasks were described as a set of jobs and then submitted to the PanDA
Server.

This approach introduced inefficiency both with usability and resource
utilization~\cite{borodin2015unified}. Ideally, users should not have to reason
in terms of jobs: Users conceive analyses in terms of one or more, possibly
related tasks; the `job' abstraction is required by the execution middleware,
i.e. PanDA. Further, a static partitioning of tasks into jobs does not take into
account the heterogeneity and dynamicity of the resources of the pilots on which
each job will be executed.

Another problem of static job sizing is that PanDA instantiates pilots on sites
with different type of resources and different models of availability of those
resources. An optimal sizing of each job should take into account these
properties. For example, sites may offer cores with different speed, networking
with different amount of bandwidth, and resources could be guaranteed to be
available for a defined amount of time or could disappear at any point in time
as with opportunistic models of resource provision.

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
the application layer submit task descriptions to JEDI
(Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
database (Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of
different size, depending on both static and dynamic information about available
resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
resources that best match jobs' requirements, and submitted to the PanDA Server
for execution (Fig.~\ref{fig:architecture}:4).

@@

Once jobs are bound to sites, the Brokerage module communicates to the Data
Service module what data sets need to be made available on what site
(Fig.~\ref{fig:architecture}:6). The Data Service communicates these
requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
takes care of  replicating them on the required sites
(Fig.~\ref{fig:architecture}:8).

Meanwhile, AutoPyFactory defines a set of suitable PanDA Pilots and concurrently
submits them to a Condor-G broker (Fig.~\ref{fig:architecture}:9). This broker
submits these pilots wrapped as jobs or VMs to the required sites
(Fig.~\ref{fig:architecture}:10).

When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
module of the PanDA Server, requesting  a job to execute
(Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
Buffer module for a job that is bound to the site of that pilot and ready to
be executed. Task Buffer checks the global queue (i.e., the PanDA database)
and, upon availability, returns a suitable job to the Job Dispatcher. The Job
Dispatcher dispatches that job to the PanDA Pilot
(Fig.~\ref{fig:architecture}:12).

Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
subprocess for the execution of the job's payload. The job is set up, input data
are transferred from the designated staging-in location
(Fig.~\ref{fig:architecture}:13), and the job's payload is executed
(Fig.~\ref{fig:architecture}:14). Once completed, output is transferred to the
staging-out location (Fig.~\ref{fig:architecture}:15).

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
    \caption{PanDA WMS architecture. Numbers indicates the execution process
    based on JEDI. Several subsystems and components, the architectural
    details of PanDA Pilot and the communication among components
    are abstracted for clarity.}
\label{fig:architecture}
\end{figure}

>>>>

% PanDA is a Workload Management System (WMS) designed to support the execution of
% distributed workloads and workflows via pilots. WMS is middleware for
% discovering and selecting resources, submitting tasks of workloads and
% workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
% abstraction that enables multi-level scheduling by decoupling resource
% acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
% implemented, a pilot is scheduled on a site and, once active, tasks are
% scheduled to the pilot, not to the site's scheduler.
%
% Pilot-capable WMS enable high throughput of tasks execution while supporting
% interoperability across multiple sites. This is particularly relevant for LHC
% experiments, where millions of tasks are executed across multiple sites every
% month, analyzing and producing petabytes of data. The design of PanDA WMS
% started in 2005 to support ATLAS. PanDA went into
% production for the LHC Run 1 on 2009, and was then extended with new subsystems
% for Run 2 in 2015.

PanDA is a Workload Management System (WMS)~\cite{marco2009glite} designed to
support the execution of distributed workloads and workflows via
pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high throughput
of tasks execution via multi-level scheduling while supporting interoperability
across multiple sites. This is particularly relevant for LHC experiments, where
millions of tasks are executed across multiple sites every month, analyzing and
producing petabytes of data. The design of PanDA WMS started in 2005 to support
ATLAS. PanDA went into production for the LHC Run 1 on 2009, and was then
extended with new subsystems for Run 2 in 2015.

@@

PanDA's application model assumes tasks grouped into workloads and workflows.
Tasks represent a set of operations performed on data sets stored in one or more
input files. Tasks are decomposed into jobs, where each job consists of the
task's operations performed on a partition of the task's data set. Since 2005, a
certain amount of parallelism has been progressively introduced for job
execution~\cite{crooks2012multi} but, so far, HEP applications have not required
MPI.
% Jobs are supposed to be capable of setting up their own execution
% environment and have a minimal set of common dependences.

PanDA's usage model assumes multitenancy of resources and at least two types of
HEP users: individual researchers and groups executing so called `production'
workflows.
% Users can submit tasks and workflows to the PanDA WMS at any point in
% time, directly or via dedicated application frameworks.
PanDA's security model is based on separation between authentication,
authorization and accounting for both single users and group of users. Both
authentication and authorization are based on digital certificates and on the
virtual organization abstraction~\cite{foster2001anatomy}.

@@

PanDA's design supports provenance and traceability for both jobs and data.
Attributes enable provenance by linking jobs and data items, providing
information like ownership or project affiliation. States enable traceability by
providing information about the stage of the execution in which each job or data
item is or has been.
% Some attributes are assumed to be immutable across execution and jobs and data
% items are assumed to be always in one and only one state.

@@

The implementation of PanDA WMS consists of several interconnected subsystems,
most of them built from off-the-shelf and Open Source components. Subsystems
communicate via dedicated API or HTTP messaging, and each subsystem is
implemented by one or more modules. Databases are used to store eventful
entities % and entities' description
like tasks, jobs and  input/output data, and to store information about sites,
resources, logs, and accounting.

@@

% The relation between tasks and jobs can be one-to-one or one-to-many, and the
% conversion between the two can by static or dynamic.

During LHC Run 1, PanDA required users to perform a static conversion between
tasks and jobs: tasks were described as a set of jobs and then submitted to the
PanDA Server. This introduced inefficiency both with usability and resource
utilization~\cite{borodin2015unified}. Ideally, users should conceive analyses
in terms of one or more, possibly related tasks, while the execution manager
(i.e., PanDA) should partition tasks into jobs, depending on execution
constraints. Further, the static partitioning of tasks into jobs does not take
into account the heterogeneity and dynamicity of the resources on which each job
will be executed, introducing inefficiencies.

Another problem of static job sizing is that PanDA instantiates pilots on sites
with different type of resources and different models of availability of those
resources. An optimal sizing of each job should take into account these
properties. For example, sites may offer cores with different speed, networking
with different amount of bandwidth, and resources could be guaranteed to be
available for a defined amount of time or could disappear at any point in time
as it happens with opportunistic models of resource provision.

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users
submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that stores
them into a queue implemented by a database (Fig.~\ref{fig:architecture}:2).
Tasks are partitioned into jobs of different size, depending on both static and
dynamic information about available resources (Fig.~\ref{fig:architecture}:3).
Jobs are bound to sites with resources that best match jobs' requirements, and
submitted to the PanDA Server for execution (Fig.~\ref{fig:architecture}:4).

@@

Once jobs are bound to sites, the Brokerage module communicates to the Data
Service module what data sets need to be made available on what site
(Fig.~\ref{fig:architecture}:6). The Data Service communicates these
requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
replicates data sets on the required sites (Fig.~\ref{fig:architecture}:8).

Meanwhile, AutoPyFactory defines PanDA Pilots, submitting them to a Condor-G
agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots wrapped
as jobs or VMs to the required sites (Fig.~\ref{fig:architecture}:10).

When a PanDA Pilot becomes available, it requests the Job Dispatcher module of
the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11). The Job
Dispatcher interrogates the Task Buffer module for a job that is bound to the
site of that pilot and ready to be executed. Task Buffer checks the global queue
(i.e., the PanDA database) and, upon availability, returns a job to the Job
Dispatcher. The Job Dispatcher dispatches that job to the PanDA Pilot
(Fig.~\ref{fig:architecture}:12).

Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
subprocess for the execution of the job's payload. Input data are transferred
from the stage-in location (Fig.~\ref{fig:architecture}:13) and the job's
payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output is
transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based execution
    process. Several subsystems, components, and architectural and
    communication details are abstracted for clarity.}
\label{fig:architecture}
\end{figure}
